---
title: "My title"
# Descriptive, informative, specific title
subtitle: "My subtitle if needed"
author: 
  - Lexi Knight
thanks: "Code and data are available at: https://github.com/LexiKnight/neighborhoods_toronto/tree/main"
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
# Tell reader top level findings
# One thing we learn about the world because of this paper?
# Abstracts need to be “tightly written”, almost terse. Remove unnecessary words. Do not include more than four sentences.

format: pdf
number-sections: true
bibliography: references.bib
---

# Introduction

I use @citeR and @rohan.

The remainder of this paper is structured as follows. @sec-data....

Notes
2-3 paragraphs of content
Additional paragraph that sets out the remainder of the paper. telegraph the rest of the paper: “Section 2…, Section 3….”.
In the introduction, please be more specific about your findings
Example
The remainder of this paper is structured as follows. First, the methods used to analyze global and local patterns of segregation are reviewed. Next, the data and study region are described, the multilevel models are detailed, and the results are presented. Finally, the contributions of this study for understanding residential income segregation are discussed, the limitations are highlighted, and directions for future research are proposed


# Data {#sec-data}

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false

ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = "none") +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Talk more about it.

And also planes (@fig-planes).


```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

analysis_data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))

analysis_data |> 
  ggplot(aes(x = width, y = length)) +
  geom_point(alpha = 0.8) +
  theme_minimal() +
  labs(x = "Wing width (mm)",
       y = "Wing length (mm)")
```

# Results

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

first_model <-
  readRDS(file = here::here("models/first_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

modelsummary::modelsummary(
  list(
    "First model" = first_model
  ),
  statistic = "mad",
  fmt = 2
)
```


# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this. 

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}


\newpage


# References


